---
title: 'Reproducible Research: Peer Assessment 1'
author: "JIN Xiaoyang"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
---
```{r global_options, echo=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo=TRUE, fig.path="figure/", dev="svg")
```

## Loading and preprocessing the data
load data and transform the second column to Date format
```{r}
if(!file.exists("activity.csv")) {
    unzip("activity.zip")
}

dat <- read.csv("activity.csv", stringsAsFactors = FALSE)
dat$date = as.Date(dat$date)
dat.clean = dat[complete.cases(dat),]
```

## What is mean total number of steps taken per day?
histogram of the total number of steps taken each day
```{r}
dat.daysum = aggregate(steps~date, data=dat.clean, sum)
ggplot(aes(steps), data=dat.daysum) +
    geom_histogram(binwidth=2500) +
    ylab("number of days")
```

mean number of steps taken per day
```{r}
mean(dat.daysum$steps)
```

median number of steps taken per day
```{r}
median(dat.daysum$steps)
```

## What is the average daily activity pattern?
the average daily activity pattern
```{r}
dat.intsum = aggregate(steps~interval ,data=dat.clean, mean)
ggplot(dat.intsum, aes(x=interval, y=steps)) + geom_line()
```

get the interval containing the maximum number of steps
```{r}
dat.intsum$interval[which.max(dat.intsum$steps)]
```

## Imputing missing values
total number of missing values in the dataset
```{r}
nrow(dat) - nrow(dat.clean)
```

filling in all of the missing values in the dataset according to the mean for that 5-minute interval
```{r}
dat.filled = dat
dat.filled[is.na(dat$steps), "steps"] = round(
    dat.intsum[match(dat[is.na(dat$steps), "interval"],
                     dat.intsum$interval),
               "steps"])
```

histogram of the total number of steps taken each day
```{r}
dat.filled.daysum = aggregate(steps~date, data=dat.filled, sum)
ggplot(aes(steps), data=dat.filled.daysum) +
    geom_histogram(binwidth=2500) +
    ylab("number of days")
```

mean number of steps taken per day
```{r}
mean(dat.filled.daysum$steps)
```

median number of steps taken per day
```{r}
median(dat.filled.daysum$steps)
```

It shows that these values differ from the estimates from the first part of the assignment. Imputing missing data makes the mean and median value smaller.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
dat.filled$weekdays = factor(format(dat.filled$date, "%u") %in% c('6', '7'),
                             labels = c("weekdays", "weekends"))
dat.filled.intsum = aggregate(steps~interval+weekdays, data=dat.filled, mean)
ggplot(dat.filled.intsum, aes(x=interval, y=steps)) + geom_line() + facet_wrap(~weekdays, nrow=2)
```

Yes, the patterns between weekdays and weekends are different.
